MC-based feasibility study of new sampling calorimeter for measuring the $\gamma$ incident angle

### 2022-01-24 paper.tex
update

\begin{abstract}
We present studies on the detector configuration to measure the incident angle of the $\gamma$, which has the energy ranging from hundred MeV to few GeV using a new sampling calorimeter. The sampling calorimeter consists of alternating 1-mm-thick lead sheets producing the electromagnetic (EM) shower and strips of 5-mm-thick plastic scintillators measuring the energy deposit by the shower particles. The strips are arranged along to the alternating horizontal and vertical directions to measure the transverse profile by combining the both directions at a given longitudinal position. In this paper, we used the GEANT4 to simulate energy deposit to the individual strips of the calorimeter, and the $\XGB$ to reconstruct the incident angle from energy deposits.

The angular resolution weakly depends on the width of the strips up to 15 mm and becomes worse with the increasing width. It is found that the incident angle is reconstructed with the compatible angular resolution using front 5 radiation length (5$X_{0}$) of the detector. The resolution does not depend on the incident angle up to 30 degrees, and largely depends on the incident energy, which can be expressed as 0.1+1.2$/ \sqrt{E_\gamma}$.

@@

\section{Motivation}
\label{sec:mot}
The EM calorimeter has played an important role in experimental studies on the nuclear and particle physics. Various materials have been developed for better energy and timing resolution so far~\cite{Calorimeter}. On the other side, the sampling calorimeter becomes popular in the large-scale high energy experiment mainly due to its cost-effectiveness. The sampling calorimeter consists of alternating passive converters that generate the EM shower and active counters that measure the energy deposit. Since the energy deposit in the passive convert is not measured, the fluctuation of energy deposits in the converter and the counter, called sampling fluctuation, determines the energy resolution of the calorimeter. The sampling fluctuation can be optimized by combining the converter and the counter with a specific portion. A 5-m-long cylindrical calorimeter made of alternating lead and plastic scintillating plates is an example of the sampling calorimeter~\cite{E391a_barrel}.

The layer structure of the sampling calorimeter enables us to measure the individual profile of the shower along to the beam direction, and the incident angle can be estimated by correlating energy deposits on neighboring layers. The measurement of the incident angle largely benefits the background rejection. Of particular, The incident angle will be an important tool for the KOTO experiment~\cite{KOTOproposal}.

Since the EM shower is evolving via stochastic processes, the incident angle would be reconstructed with a certain angular resolution. Energy deposits from random processes are used to reconstruct the incident angle with the machine learning, which provides better angular resolution. The evolution of the EM shower is generated with the GEANT4, and $\XGB$ was used to reconstruct the incident angle.

In sec~\ref{sec:ems}, generic explanation of the EM shower and its evolution. We describe the reconstruction and angular resolution of the sampling calorimeter in sec~\ref{sec:res}. We summarize the study in sec~\ref{sec:sum}.

\section{ELECTROMAGNETIC SHOWER}
\label{sec:ems}
We started the study with a detector setup consisting of alternating lead sheet as the passive converter and strips of the plastic scintillator as the active counter. The thickness of lead is 1 mm and scintillator is 5 mm, respectively. The lead sheet has a cross-section of  500 mm $\times$ 500 mm and each scintillator strips of 500 mm $\times$ 15 mm  being arranged along the X- and Y-direction alternatively along Z-direction as shown in  Fig.~\ref{fig:det_conf}. Total number of alternating layers is 105 which corresponds to 20 radiation length (20 Xo) and possible to contain all of the shower particles generated by $\gamma$ in the region of medium energy (from  100 MeV to few GeV).

\begin{figure}[!hbt]
\includegraphics[width=0.85\textwidth]{figures/Sec2/Prototype_samplingcal.pdf}

\caption{ The first trial of detector setup for simulation study on angle measurement. It is a sampling calorimeter consisting of alternating lead sheets and strips of plastic scintillator. }
\label{fig:det_conf}
\end{figure}

The generation of incident $\gamma$ and its interaction with detector materials was simulated by using the GEANT4 package (ver. xxx) with the standard EM sub-packages~\cite{GEANT4}.
The $\gamma$ entered the center of the detector and moved to the positive direction of Z-axis, where the surface of the first layer is defined as a starting point (Z=0). Its incident angle is defined as the polar angle ($\theta$) along to the Z-direction, that is, $\theta = 0$ for the $\gamma$ enters perpendicular to the detector surface. %The generated energies are in range from hundred MeV to few GeV which is interested to the KOTO experiment. 
Figure~\ref{fig:Evt_Dis} shows a typical event display describing the energy deposit to the each scintillator strips in both of X-Z and Y-Z plane for 1-GeV $\gamma$ entering in perpendicular to the center of the detector ($\theta=0$ and $\phi=0$).

\begin{figure}[!hbt]
\includegraphics[width=0.45\textwidth]{figures/EventDisplay.jpg}
\includegraphics[width=0.45\textwidth]{figures/EventDisplay.jpg}
\caption{Distribution of energy deposit to the each scintillator strips in X-Z plane (Left) and Y-Z plane (Right) for 1-GeV $\gamma$ entering perpendicularly to the detector ($\theta=0$ and $\phi=0$).}
\label{fig:Evt_Dis}
\end{figure}

The GRANT4 can provide reliable results on the electromagnetic shower produced by the incident $\gamma$. Its longitudinal profile depends on incident energy and can be expressed \cite{Longo} by the three parameters as following;
\begin{equation}
    F(E,t)=A  t^\alpha e^{-bt},
\end{equation}
where t is the position of scintillator strips along the Z-direction. As shown in Fig.~\ref{fig:Longitudinal}, the Eq. 1 describes the obtained average energy deposit on each layer of scintillator strips reasonably. In the case of the sampling calorimeter, however, the energy deposit on the converters and counters is caused by multiple processes and its fraction is changed along with the shower develpoment.

As a result, the ratio of the energy deposit in counter and in converter changes according to the layer position as shown in Fig.~\ref{fig:Longitudinal} (right). In order to describe the shower profile in the sampling calorimeter, we need to take into account the layer dependent of the energy deposit. Since we will use the simulated results only for current study on angle measurement, we did not proceed any further analytic study on the process.

\begin{figure}[!hbt]
\includegraphics[width=0.48\textwidth]{figures/LongProfile.jpg}
\includegraphics[width=0.48\textwidth]{figures/LayerVisibleRatio.jpg}+
\caption{Generated average energy deposit on the scintillator strips along Z axis(left) and visible ratio for each layer.}
\label{fig:Longitudinal}
\end{figure}

On the other hand, the lateral spread of the shower
%is interpreted as a distribution of energy deposit in crossing positions of the X- and the Y-strips in neighbouring two layers.  This spread 
is expressed as the the Molière radius defined ~\cite{PDG} as 
\begin{equation}
R_M= 21.2~{\rm MeV}\cdot X_0 /E_c,
\end{equation}
where $X_0$ is the radiation length and $E_c$ is the critical energy of the detector, respectively. 
Since the sampling calorimeter contains different materials placed at the different localized positions, we defined the $R_M$ as the radius of cylinder containing 90\% of the energy deposit instead of using the Eq. (2). The left panel of Figure~\ref{fig:Moliere_vis} shows energy deposit per each layer (black line) and integrated energy deposit up to the layers. We can deduce the Molière radius as 70 mm, which is largely depends on the position of its measurement. The right panel of Figure~\ref{fig:Moliere_vis} shows the radius as a function of scintillator layer, which increases along the shower depth. The first three bins show large radius because of large contribution of shower particles generated opposite to the incident direction.

\begin{figure}[!hbt]
\includegraphics[width=0.44\textwidth]{figures/Edep_R.jpg}
\includegraphics[width=0.48\textwidth]{figures/Moliere_layer.jpg}
\caption{Left:Energy deposit for given radius as a differential(black line) and integral distribution(red line) Right: Moliere radius as a function of ID of scintillator-strip layers.}
\label{fig:Moliere_vis}
\end{figure}

With the detector configuration shown in Fig.~\ref{fig:det_conf}, we can get three dimensional information on the showers. By using the obtained data as an input of the machine learning package, we can extract the incident angle of the $\gamma$ as explained following section. 
%When we apply the three dimensional information on the generated shower profile, we can reconstruct incident $\gamma$ more correctly by integrating the energy deposit. 
%For simplicity in numerical calculation, %For the sampling calorimeter consisting of different materials, 
%let us define the Molière radius as the radius of cylinder containing 90\% of shower energy. As shown in Fig.~\ref{fig:Moliere_vis}, the configuration of 1-mm lead and 5-mm scintillator shows $\rho_M=70$ mm in average which changes with the shower depth. %That is, we can expect an improvement on separating two gammas closely incident to the calorimeter. We will study on this subject and not discuss any more in this paper.
%the For the detector consisting of 1mm lead and 5mm plastic scintillator contain 90\% of energy deposit inside the cylinder of radius of $\rho_M=70$ mm. %It had better to compare with CsI crystal

%The energy resolution of the sampling calorimeter is mainly determined by the fluctuation of the visible ratio defined as 
%\begin{equation}
%R = { \frac{E_{act}} {E_{act}+E_{pas}}},
%\end{equation}
%where $E_{act}$ and $E_{pas}$ are energy deposit to the active and passive part of the calorimeter, respectively. This visible ratio is roughly given by the ratio of material budget between the converter and the counter. 
%is simply given by the fraction of the active component to that of the passive component, and its fluctuation becomes large when the ratio is small.
%by thickness of the converter. It is because the energy deposit at the active component is done through not only passage of the electron-positron pair but also low energy process such as compton scattering not reach to the active counter. 
%As shown in Fig.~\ref{fig:energy_res}, the mean of the visible ratio for the configuration of 1-mm lead and 5-mm scintillator is  0.35 and its energy resolution is expected 4\% for 1 GeV photon. However, the visible ratio can be improved with the same amount of composed materials when we use thinner layers. Figure YYY shows the change of the visible ratio for the same ratio $R=t_{lead}/t_{scin}$=0.2 according to the thickness of the converter (thinner scintillator too). When we make the lead sheet thinner as a factor of 5, the visible ratio improves as XXX. This is because the number of low energy electrons and positrons penetrating the lead sheet is increased in case of thinner converter. As a result of the increasing probability, the sampling fluctuation becomes small and the energy resolution improved as a result.
%This energy resolution is improved as a result thinner converter and counter keeping the visible ratio constant.

%\begin{figure}[!hbt]
%\includegraphics[width=0.48\textwidth]{figures/3.2/eres_thickness.pdf}
%\includegraphics[width=0.48\textwidth]{figures/3.2/energy_resolution.pdf}
%\caption{Visible ratio of the sampling calorimeter for different configuration (Left) and energy resolution as a function of incident $\gamma$ energy (right).}
%\label{fig:energy_res}
%\end{figure}

%\input{Sec2.tex}
%\section{method}
%\label{sec:ana}

%\subsection{Geant4 simulation}
%In this paper, the photon shower and corresponding detector responses, the prerequisites for the machine learning inputs, are simulated based on the Geant4 framework. Among several physics packages, the $\rm{FTFP\_BERT}$ physics package is used for electromagnetic and hadronic interactions as it is generally recommended for the high energy physics calorimetry.

%The materials used in detector construction are selected from the Geant4 NIST material database: $\rm{G4\_Pb}$ for the lead and $\rm{G4\_PLASTIC\_SC\_VINYLTOLUENE}$ for the scintillator material.

%\subsection{Detector construction}
%We constructed a prototype sampling electromagnetic calorimeter with alternative stacking of thin Pb sheets and scintillator strips. A Pb sheet has a dimension of 50 cm (w) $\times$ 50 cm (l) $\times$ 1 mm (t). And a single layer of scintillator plate consists of 25 long scintillator segments, and each segment has a dimension of 2 cm (w) $\times$ 50 cm (l) $\times$ 5 mm (t). The full geometry consists of 105 layers  which is corresponding to 20 $X_0$. In order to investigate the detector properties and test the reconstruction performance of the indicent angle of $\gamma$, we generated $\gamma$ beam data samples with several incident angles and energies.
%In this study, we utilized a physics package $FPFT\_BERT$ for the hadronic and electromagnetic interactions
%We started the study with a setup of sampling calorimeter consisting of alternating lead sheet and strips of plastic scintillator as shown in Fig.~\ref{fig:det_conf}. The thickness of lead is 1 mm and scintillator is 5 mm, respectively. The lead sheet has a cross-section of  500 mm $\times$ 500 mm and the scintillator strip of 500 mm $\times$ 15 mm. By arranging the strips in X- and Y-direction alternatively along Z-direction. The generated photon moves along with Z-direction and its incident angle was defined as a polar angle to the direction.  With the configuration, we can get shower profiles in X-Z and Y-Z planes in turn, and estimate the incident angle from them. The number of alternating layers is 105, corresponding to 20 radiation length (Xo), in order to contain shower particles fully.

%\begin{figure}[!hbt]
%%\includegraphics[width=0.85\textwidth]{figures/Sec2/ang_conf.pdf}
%\includegraphics[width=0.85\textwidth]{figures/Sec2/Prototype_samplingcal.pdf}

%\caption{ The first trial of detector setup for simulation study on angle measurement. It is a sampling calorimeter consisting of alternating lead sheets and strips of plastic scintillator. }
%\label{fig:det_conf}
%\end{figure}

%\subsection{Electromagnetic shower Generation}
%\subsection{Properties of electromagnetic shower (EM shower)}
%We generated the EM shower generated inside the detector with GEANT4, which provides reliable results for them polish-up later).

%\subsection{Properties of electromagnetic shower (EM shower)}

%In general, the less the photon shower develops in the transverse direction, the more separable the photon clusters, and the higher the visible ratio, the better the energy resolution. However, these properties are incompatible as they have strong inverse correlation. For the trade-off configuration, we investigated the Molière radius and the visible ratio with several configurations.

% 어떤 configuration, energy? 등등
% 2.5, 5, 10, 15, 20 mm for scint., 0.5, 1, 2 mm for lead 1GeV gamma
% For energy resolution, 100 MeV step 100MeV ~ 2GeV gamma

%\subsection{Backsplash}
% backsplash particle이 veto signal을 만들어 accidental event loss를 만든다.
% incident photon의 에너지와 각도를 바꿔가며 front surface로 빠져나오는 backsplash particle의 개수를 카운팅
% 1개 이상의 backsplash particle이 있으면 veto에 counting

%\subsection{XGBOOST: A toolkit for machine learning}
%\label{sec:anaML}
%Among many packages for machine learning in the high energy physics area~\cite{ATLAS:2020iwa}, we used the $\XGB$, supporting a scalable tree boosting system~\cite{xgboost:2016}. The $\XGB$ package is used to reconstruct the incident angle of incident $\gamma$ particles with deposit energies in the each channel, where secondary particles of the EM shower raise signals. Input data for the $\XGB$ are deposit energies for whole channels including zero-energies to tally with the requirement that the number of input data should be identical. For the $\XGB$ to reconstruct the angle, the training procedure for the $\XGB$ needs to be preceded. The training procedure helps the $\XGB$ study how $\gamma$ interacts with the detector and deposits the energy to the each channel with respect to the incident angle for a given $\gamma$ energy. Simulations are done with the uniform incident polar angle ($\theta$) distribution from 0 to 50 degree and the uniform azimuthal angle ($\varphi$) from 0 to 360 degree for the training, which makes the $\XGB$ possible to reconstruct events having 0~$<\theta<$~50~degree and 0~$<\varphi<$~360~degree. The reconstruction of $\theta$ by the $\XGB$, completing the training procedure, is tested with the fixed $\theta$ and uniform $\varphi$ from 0 to 360 degree.

@@

The incident angle of the $\gamma$ can be estimated as a results of machine learning by using the $\XGB$, which is one of the popular packages supporting a scalable tree boosting system~\cite{xgboost:2016}. We used energy deposits on each scintillator strips as input data for the $\XGB$, and the packages connects the relationship among the input data to estimate the incident angle. ** a little detail one or two senteces***

%are deposit energies on each scintillator strip by the generated shower particles. It includes the strips of zero-energy deposit in order to satisfy the requirement that the number of input data should be identical. As the training procedure, the $\XGB$ studies profiles of energy deposit to the individual strips with respect to the incident angle for a given $\gamma$ energy.  ({\it This part should be updated)}

For the training of the $\XGB$, we generated $\gamma$'s uniformly in the range 0 to 50 degrees of polar angle ($\theta$) and  0 to 360 degrees of azimuthal angle ($\varphi$) for given energy (training sample). By using another data set (testing sample) which was generated at a fixed polar angle, we estimated performance of angular reconstruction of the detector. The number of generated $\gamma$ was
$2\times10^5$ for the training sample and $1\times10^5$ for the testing sample, respectively.

%\subsection{Machine learning configuration}
%The configuration of the $\XGB$ can be organized with five parameters, described in Tab.~\ref{tab:XgbPar}. The parameters are selected as one providing better angular resolution. It is found that the Max. depth mostly affects the angular resolution, and the resolution is not enhanced for Max. depth~$>$~15 any more. It is also worth mentioning that configurations with N\_estimators larger than 1000 do not provide better resolution with the Max. depth~=~15. The angular resolution of the reconstruction is defined as the standard deviation of Gaussian function, whose parameters are evaluated by fitting the reconstructed angle distribution. Figure~\ref{fig:angle_reco_def} shows the reconstructed $\theta$ distribution with respect to each incident $\theta$ with described training parameters. The angular resolution for all $\theta$ is estimated to be 1.3~degree.

%\begin{figure}[!hbt]
%\includegraphics[width=0.89\textwidth]{figures/Fig1_reco_def.pdf}
%\caption{ (Color online) Reconstructed $\theta$ with respect to each true incident $\theta$ dotted lines denote each Gaussian function fitting the reconstructed $\theta$ distribution. The standard deviation of all Gaussian functions are estimated to be 1.3 degree.}
%\label{fig:angle_reco_def}
%\end{figure}

%\begin{table}[hbt!]
%\centering
%\caption{Parameterization of the $\XGB$}
%\begin{tabular}{cccc}
%\hline 
%Parameter & Function & Default value & Used value \\ \hline 
%N\_estimators & The number of decision trees & N.A. & 1000 \\  
%Max. depth & Possible maximum depth of tree structure & 6 & 15 \\ 
%Subsample & Fraction of total data used for a single decision & 1 & 1 \\ 
%Learning rate & Step length for calculation & 0.3 & 0.08 \\ 
%Gamma & Requirement on minimum loss function & 0 & 0 \\ 
%\hline
%\end{tabular}
%\label{tab:XgbPar}
%\end{table}

%\subsection{Detector dimension}
%\subsection{Angle reconstruction using Machine Learning}

%With the results of the $\XGB$ by using the training sample generated with the uniform distributions in both of polar and azimuthal angles, we estimated the incident angle of photons generated with a fixed value for a given energy. 
Figure~\ref{fig:angle_10degree} shows the distribution of reconstructed angles for 1-GeV $\gamma$ entered detector with 10 degrees of incident angle. %The Gaussian distribution (black solid line) provides the proper mean value,  the $\XGB$ can provide correct answer. 
We can get proper mean value of the distribution, however, it has heavier tail compared to that of the Gaussian distribution. In order to describe the obtained result more correctly, we tried to fit it with the Generalized Gaussian (GG) function.

\begin{figure}[!hbt]
\includegraphics[width=0.7\textwidth]{figures/GG_fit.jpg}
\caption{ (Color online) Reconstructed $\theta$ with respect to 10 degrees of incident $\theta$. The gaussian function can not describe the tail region (red dotted line). The generalized gaussian function (black solid line) shows much better agreement with the obtained result.}
\label{fig:angle_10degree}
\end{figure}

%In order to include the heavier tail more correctly in the performance study, we tried to fit the distribution with two different function; one is a convolution of a Lorentz and Gaussian distribution (Voigt function) and the other is a Generalized Gaussian function.
%The Voigt fuction expressed as 
%\begin{equation}
%V(x; \sigma, \gamma) = \int_{- \infty}^{\infty} G(x';\sigma)L(x-x';\gamma) dx',
%\end{equation}
%becomes to be Lorentz distribution $L(x; \gamma)$ in case of $\gamma$=0 
%\begin{equation}
%L(x;\mu, \gamma) = {\gamma \over {\pi ((x-\mu)^2 + \gamma^2)}}
%\end{equation}
%of peak position at x=$\mu$ in case of $\sigma$ =0 
%and to be Gaussian distribution $G(x;\sigma)$
%\begin{equation}
%G(x;\mu, \sigma) = {1 \over {\sqrt{2 \pi} \sigma}} e^{-(x-\mu)^2/2\sigma^2} 
%\end{equation}
%when the $\gamma$ equal to be zero. The function is available in the analysis framework of the ROOT~\cite{Voigt}, and  improves the fitness a little.

The GG function provides more flexible modeling to data than the Gaussian can do by introducing an additional parameter to describe the shape of distribution. This function was suggested by Subbotin~\cite{Subbotin} in 1923 treating the accidental errors not only to be continuous but also to be expanded in the power series. Its probability density function is expressed as 
%has received considerable attention especially at the image processing studies\cite{Dytso},  
%Among various parametrization of the function, we take that its probability density distribution (pdd) is given as~\cite{Gonzalez}  
\begin{equation}    
  % GG(x;\mu, \sigma,     p) = {C_p \over \sigma} e^{-{|x-\mu|^p}/{2 \sigma ^p}}, 
   %f(x; \mu, \alpha, \beta) = {\beta \over {2 \alpha \Gamma(1/\beta)}} e^{-(|x-\mu|/\alpha)^\beta} , 
     f(x; \mu, \alpha, \beta) = \frac{\beta}{2 \alpha \Gamma(1/\beta)} e^{-(|x-\mu|/\alpha)^\beta} , 
%   f(x; \mu, \alpha, \beta) = {\beta \over {2 \alpha \Gamma (1/\beta)}} e^{-(|x-\mu|/\alpha)^\beta} , 
\end{equation}    
where $\beta$ is the so-called shaping factor
%Also  
%\begin{equation}    
%    A ={ {\beta p}  \over {2\Gamma(1/p)}},
%\end{equation}
and the $\Gamma(.)$ is the gamma function.
It becomes the Laplace distribution if $\beta$=1 
\begin{equation}
LP(x;\mu, \sigma) = \frac{1}{\sqrt{2}\sigma} e^{-{\sqrt{2}{|x-\mu|} / \sigma}} 
\end{equation}
and Gaussian distribution
\begin{equation}
G(x;\mu, \sigma) = \frac{1}{\sqrt{2 \pi} \sigma} e^{-(x-\mu)^2/2\sigma^2} 
\end{equation}
for $\beta$=2.  The second moment of the GG function is given as
\begin{equation}    
    E[X]^2  = \alpha^2 \frac{\Gamma(3/p)}{\Gamma(1/p)}  \equiv \sigma^2 , 
\end{equation}
which becomes the variance of the Gaussian distribution in case of the $\beta$=2. This value can be interpreted equal footing regardless of the $\beta$-values, and will be used as the angular resolution for given detector configuration from now.
%we can treat it as a width of the distribution which contains same meaning %regardless of the p-values.

>>>>

\begin{abstract}
We present studies on the detector configuration to measure the incident angle of the $\gamma$ with energies ranging from hundred MeV to few GeV using a new sampling calorimeter. The sampling calorimeter consists of alternating 1-mm-thick lead plates producing the electromagnetic (EM) shower and strips of 5-mm-thick plastic scintillators measuring the energy deposit by the shower particles. The strips are arranged along the alternating horizontal and vertical directions to measure the transverse profile by combining both directions at a given longitudinal position. In this paper, we used the GEANT4 to simulate energy deposit to the individual strips of the calorimeter, and the $\XGB$ to reconstruct the incident angle from energy deposits.

The angular resolution weakly depends on the width of the strips up to 15 mm and becomes worse with the increasing width. It is found that the incident angle is reconstructed with the compatible angular resolution using front 5 radiation length (5$X_{0}$) of the detector. The resolution does not depend on the incident angle up to 30 degrees, and largely depends on the incident energy, which can be expressed as 0.2+1.1$/ \sqrt{E_\gamma}$.

@@

\section{Motivation}
\label{sec:mot}
The EM calorimeter has played an important role in experimental studies for the nuclear and particle physics. Various materials have been developed for better energy and timing resolution so far~\cite{Calorimeter}. On the other side, the sampling calorimeter becomes popular in the large-scale high energy experiment mainly due to its cost-effectiveness. The sampling calorimeter consists of alternating passive converters that generate the EM shower and active counters that measure the energy deposit. Since the energy deposit in the passive convert is not measured, the fluctuation of energy deposits in the converter and the counter, called sampling fluctuation, determines the energy resolution of the calorimeter. The sampling fluctuation can be optimized by combining the converter and the counter with a specific portion. A 5-m-long cylindrical calorimeter made of alternating lead plates and plastic scintillating plates is an example of the sampling calorimeter~\cite{E391a_barrel}.

The layer structure of the sampling calorimeter allow one to measure the individual profile of the shower along the beam direction, and the incident angle can be estimated by correlating energy deposits on neighboring layers. The measurement of the incident angle largely benefits the background rejection. Of particular, The incident angle will be an important tool for the KOTO experiment~\cite{KOTOproposal}.

Since the EM shower is evolving via stochastic processes, the incident angle would be reconstructed with a certain angular resolution. Energy deposits from random processes are used to reconstruct the incident angle with the machine learning, which provides better angular resolution. The evolution of the EM shower is generated with the GEANT4, and the $\XGB$ was used to reconstruct the incident angle.

In sec~\ref{sec:ems}, generic explanation of the EM shower and its evolution are described. We describe the reconstruction and angular resolution of the sampling calorimeter in sec~\ref{sec:res}. We summarize the study in sec~\ref{sec:sum}.

\section{ELECTROMAGNETIC SHOWER}
\label{sec:ems}

\begin{figure}[!hbt]
\includegraphics[width=0.85\textwidth]{figures/Sec2/Prototype_samplingcal.pdf}
\caption{ \textcolor{red}{plots should be updated} Schematic view of the sampling calorimeter. It consists of alternating lead plates and scintillator strips.  }
\label{fig:det_conf}
\end{figure}

The sampling calorimeter was designed as alternating 1-mm-thick lead plates for passive converters and 5-mm-thick strips of the plastic scintillator for active counters. The cross section of the lead plate and the scintillator is 525 mm $\times$ 525 mm and 15 mm $\times$ 525 mm, respectively. Strips are alternatively aligned along $x$ and $y$ direction as shown in Fig.~\ref{fig:det_conf}. The total number of alternating layers is 105, which corresponds to 20$X_{0}$ to make it possible to absorb whole products from the EM shower ranging from 100 MeV to few GeV energy.

\begin{figure}[!hbt]
%\includegraphics[width=0.45\textwidth]{figures/EventDisplay.jpg}
%\includegraphics[width=0.45\textwidth]{figures/EventDisplay.jpg}
\includegraphics[width=0.48\textwidth]{figures/SingleEventXZHit.pdf}
\includegraphics[width=0.48\textwidth]{figures/SingleEventYZHit.pdf}
\caption{ An event display of the energy deposit to each scintillator strip in the $x$--$z$ plane (left) and the $y$--$z$ plane (right) for the 1~GeV $\gamma$ perpendicularly entering to the detector ($\theta=$~0).}
\label{fig:Evt_Dis}
\end{figure}

The interaction of incident $\gamma$ with the detector was simulated using the GEANT4 (ver. 4.10.06) with the standard EM sub-packages~\cite{GEANT4}. The surface of the detector is defined as $z=$~0, and the $\gamma$ starts to interact with the detector. the incident angle of the $\gamma$ is defined as the polar angle ($\theta$) along the $z$ direction, where $\theta = 0$ case denotes that the $\gamma$ perpendicularly enters into the detector surface. Figure~\ref{fig:Evt_Dis} shows the event display illustrating the energy deposit on the each scintillator strip in both of $x$--$z$ and $y$--$z$ planes for the 1~GeV~$\gamma$ with $\theta = 0$.

%\begin{figure}[!hbt]
%\includegraphics[width=0.48\textwidth]{figures/LayerVisibleRatio.jpg}
%\includegraphics[width=0.48\textwidth]{figures/Moliere_layer.jpg}
%\caption{  \textcolor{red}{plots should be updated} Sampling fraction (left) and Moliere radius (right) from 1 GeV %$\gamma$ events ($\theta=0$) for each layer}
%\label{fig:sc_prop}
%\end{figure}

%Figure~\ref{fig:sc_prop} shows the sampling fraction and the Moliere radius~\cite{PDG} for each layer with 1~GeV $\gamma$. In this paper, Moliere radius is defined as the distance including 90\% of (deposited) energy from the center of energy. The visible ratio (Moliere radius) decreases (increases) with increasing layer ID. It is, therefore, estimated to have better energy resolution and position separation with front layers. Note that the very front few layers are largely affected by secondaries going backward.

@@

The incident angle of the $\gamma$ is reconstructed with the $\XGB$, which is one of the popular machine learning toolkit providing a scalable tree boosting system~\cite{xgboost:2016}. The machine learning is required to be trained. The feature, characteristic of a phenomenon, and the target, output to be predicted, are essential inputs for the training. The machine learning correlates each feature with the corresponding target for different features. Trained machine learning takes measured features and gives the target based on correlations which are made during the training. In this paper, the feature corresponds to energy deposits on each scintillator strip and the target corresponds to the incident angle of the $\gamma$.

If the feature for the training is biased, the prediction of the $\XGB$ would be correspondingly biased as the $\XGB$ reliably believes there are no biases in all features given by a user. To minimize this, the incident angle is uniformly generated in the range 0 to 50 degrees for incident polar angle ($\theta$) and 0 to 360 degrees for incident azimuthal angle ($\varphi$), which makes the $\XGB$ train all features we are interested in. The number of samples for the training is $2\times10^5$ considering computing resources. To test the reconstruction of the incident angle, the incident angle is generated with the fixed $\theta$. Note that the incident energy is known.

\begin{figure}[!hbt]
\includegraphics[width=0.7\textwidth]{figures/GG_fit.jpg}
\caption{ Reconstructed $\theta$ distribution for $\theta=$~10~deg events. The distribution is fitted with the Gaussian function and the Generalized Gaussian function. The Generalized Gaussian function provides better description for both tails.}
\label{fig:angle_10degree}
\end{figure}

Figure~\ref{fig:angle_10degree} shows reconstructed $\theta$ for 1~GeV $\gamma$ with $\theta=$~10~degrees. The distribution is fitted with the Gaussian function and the Generalized Gaussian (GG) function. GG function provides better description for both tails of the distribution. The GG function can be expressed as
\begin{equation} 
f(x; \mu, \alpha, \beta) = \frac{\beta}{2 \alpha \Gamma(1/\beta)}e^{-(|x-\mu|/\alpha)^\beta}
\end{equation}
The variation of the GG function can be expressed as, then, $\sigma^2 \equiv \alpha^2 \Gamma(3/\beta) / \Gamma(1/\beta)$. The angular resolution of the incident angle reconstruction is defined as $\sigma$.

\begin{figure}[!hbt]
\includegraphics[width=0.89\textwidth]{figures/Fig1_reco_def.pdf}
\caption{\textcolor{red}{plots should be updated} The angular resolution for the varying AAA and the varying BBB. The best angular resolution is obtained with AAA=XXX and BBB=YYY.  }
\label{fig:par_scan}
\end{figure}

As the reconstruction of the incident angle depends on hyperparameters of the $\XGB$, which controls the details of training processes, dedicated tests to optimize the hyperparameter were executed. It is assumed that a set of hyperperameters providing the best angular resolution would be optimized. The test scans the evaluated angular resolution with different hyperparameters. Figure~\ref{fig:par_scan} shows the result of the test for AAA and BBB. The AAA works for *******, and the BBB works for *****. The AAA and BBB are determined to be XXX and YYY, respectively. Different tests determine the set of hyperparameters, which is shown in Tab~\ref{tab:XgbPar}.

\begin{table}[hbt!]
\centering
\caption{Hyperparameters of the $\XGB$}
\begin{tabular}{cccc}
\hline 
Parameter & Function & Default value & Used value \\ \hline 
N\_estimators & The number of decision trees & N.A. & 1000 \\  
Max. depth & Possible maximum depth of tree structure & 6 & 100 \\ 
Subsample & Fraction of total data used for a single decision & 1 & 1 \\ 
Learning rate & Step length for calculation & 0.3 & 0.08 \\ 
Gamma & Requirement on minimum loss function & 0 & 0 \\ 
\hline
\end{tabular}
\label{tab:XgbPar}
\end{table}

\begin{figure}[!hbt]
\includegraphics[width=0.58\textwidth]{figures/res_width.jpg}
\caption{ The angular resolution as a function of the scintillator strip width for 1~GeV $\gamma$ with $\theta=$~10~deg. The angular resolution is optimized with 15- to 20-mm-wide strips.}
\label{fig:angle_reco_width}
\end{figure}

The angular resolution was evaluated with the width of scintillator strips varying from 5 mm to 35 mm. Figure~\ref{fig:angle_reco_width} shows the angular resolution as a function of the width. 15-mm-wide strips provide the best angular resolution. The width longer than 15 mm suffers from the dilution of the EM shower, resulting in worse resolution. On the other hand, the width shorter than 15 mm results in larger size of features, which causes a negative influence on the machine learning. Dedicated study is described in fig~\ref{fig:angle_reco_width}.

\begin{figure}[!hbt]
\includegraphics[width=0.48\textwidth]{figures/Fig5_reco_graph.pdf}
\includegraphics[width=0.48\textwidth]{figures/Fig5_reco_inc.pdf}
\caption{ The angular resolution as a function of the $\gamma$ energy (left), and $p_{1}$ for 1 GeV as a function of the incident angle (right). $p_{1}$ is consistent down to $\theta=$~15~deg and changing at $\theta=$~10~deg, which is coming from the definition of $\theta$ }
\label{fig:angle_reco_dep_gr}
\end{figure}

Figure~\ref{fig:angle_reco_dep_gr} shows the angular resolution as a function of the incident $\gamma$ energy for $\theta=$~10~degrees on the left panel. The resolution is fitted with $p_{0} + p_{1}/\sqrt{E_{\gamma}(GeV)}$, where the $p_{0}$ denotes the energy-independent contribution and is estimated to be 0.2, and the $p_{1}$ denotes the energy-dependent contribution, mainly related with the development of the EM shower. The estimated $p_{1}$ for different $\theta$ can be seen on the right panel of fig.~\ref{fig:angle_reco_dep_gr}. They are deviating within $\pm$~0.05~deg range from the 1.1~deg.

\begin{figure}[!hbt]
\includegraphics[width=0.48\textwidth]{figures/res_Nlayer.jpg}
\includegraphics[width=0.48\textwidth]{figures/resol_Nlayer.jpg}
\caption{ Reconstructed $\theta$ for 1 GeV $\gamma$ with $\theta=$~10~deg with (red) and without (black) the total energy ($E_{\rm{tot}}$) selection (left), and the angular resolution as a function of front layer depth used for the reconstruction (right). The inefficiency for 4.6$X_{0}$ is estimated to be 5.8\% with 10~MeV selection. The large error bars less than 4.6$X_{0}$ is due to insufficient fitting because the reconstructed angular distributions deviated from the General Gaussian function.}
\label{fig:angle_reco_layer}
\end{figure}

The reconstruction of the incident angle is tested with the front layer instead of the full layers considering the cost-effectiveness from the fact that 99\% of the $\gamma$ generates the EM shower in front of 5$X_{0}$. The left panel of Figure~\ref{fig:angle_reco_layer} shows reconstructed angle using the front 24 layers of the detector, which corresponds to the 4.6$X_{0}$ for 1~GeV $\gamma$. Some $\gamma$s are failed to be reconstructed due to the lack of the channel having the energy deposit. The failed events are represented as a delta function near 0, and such events are removed by requiring the total energy deposit to be larger than 10~MeV. The angular resolution with the front layer is estimated to be XXX while possessing inefficiency estimated to be 5.8\%.

\begin{figure}[!hbt]
\includegraphics[width=0.6\textwidth]{figures/layer-event.jpg}
\caption{  \textcolor{red}{plots should be updated} The angular resolution for 1 GeV $\gamma$ with $\theta=$~10~deg as a function of the number of training samples with different strip widths using the first 24 layers or the full layers. }
\label{fig:multi-parameter}
\end{figure}
